---
title: "Assignment 4"
author: "Chelsea Nowlin"
date: "7/3/2020"
output: html_document
---

Q. 3 We now review k-fold cross-validation.

a. Explain how k-fold cross-validation is implemented.
  - k cross validation resamples and splits the observations into a number of groups that the data is to be split into with a single paramter (k) referencing the split groups.
  
b. What are the advantages and disadvantages of k-fold cross-validation relative to:

  i. The validation set approach ?
    - The validation approach test error rates can be highly variable among the training sets and only a subset of observations, not most or all, are used in fitting the model which tend to perform worse when training sets on fewer observations. 
  ii. LOOCV ?
    - A type of k cross validation, and computationally expensive model to run since k = n and are fitting the model equal to the number of observations, which can be a large number. Furthermore, the LOOCV has higher variance than k cross validation.
    
Q5. In Chapter 4, we used logisitc regression to predict the probability of “default” using “income” and “balance” on the “Default” data set. We will now estimate the test error of this logistic regression model using the validation set approach. Do not forget to set a random seed before beginning your analysis.

a. Fit a logistic regression model that uses “income” and “balance” to predict “default”.
```{r}
library(ISLR)
attach(Default)
set.seed(1)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(fit.glm)
```

b. Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:
  i. Split the sample set into a training set and a validation set.
```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
```
  
  ii. Fit a multiple logistic regression model using only the training observations.
```{r}
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
summary(fit.glm)
```
  
  iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the “default” category if the posterior probability is greater than 0.5.
```{r}
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
```
  
  iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.
```{r}
mean(pred.glm != Default[-train, ]$default)
```
  
b. There is a 2.54% validation set test error rate.

c. Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.
```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```
The results for test error rate varied with each split of observations. 

d. Now consider a logistic regression model that predicts the probability of “default” using “income”, “balance”, and a dummy variable for “student”. Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for “student” leads to a reduction in the test error rate.
```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance + student, data = Default, family = "binomial", subset = train)
pred.glm <- rep("No", length(probs))
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```
Adding the dummy variable had little effect in reducting the test error rate.

Q6. We continue to consider the use of a logistic regression model to predict the probability of “default” using “income” and “balance” on the “Default” data set. In particular, we will now computes estimates for the standard errors of the “income” and “balance” logistic regression coefficients in two different ways : (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the glm() function. Do not forget to set a random seed before beginning your analysis.

a. Using the summary() and glm() functions, determine the estimated standard errors for the coefficients associated with “income” and “balance” in a multiple logistic regression model that uses both predictors.

```{r}
set.seed(1)
attach(Default)
```


```{r}
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(fit.glm)
```
a. The glm standard errors for the intercept, income, and balance are
4.348e-01,
4.985e-06, and
2.274e-04 respectively.

b. Write a function, boot.fn(), that takes as input the “Default” data set as well as an index of the observations, and that outputs the coefficient estimates for “income” and “balance” in the multiple logistic regression model.
```{r}
boot.fn <- function(data, index) {
    fit <- glm(default ~ income + balance, data = data, family = "binomial", subset = index)
    return (coef(fit))
}
```

c. Use the boot() function together with your boot.fn() function to estimate the standard errors of the logistic regression coefficients for “income” and “balance”.
```{r}
library(boot)
boot(Default, boot.fn, 1000)
```
c. The bootstrap standard errors for the intercept, income, and balance are
4.344e-01,
4.866e-06, and
2.298e-04 respectively.

d. Comment on the estimated standard errors obtained using the glm() function and using your bootstrap function.

The standard errors for both methods are very similar.

Q9. We will now consider the “Boston” housing data set, from the “MASS” library.

a. Based on this data set, provide an estimate for the population mean of “medv”. Call this estimate μ^.
```{r}
library(MASS)
attach(Boston)
mu.hat <- mean(medv)
mu.hat
```

b. Provide an estimate of the standard error of μ^. Interpret this result.
```{r}
se.hat <- sd(medv) / sqrt(dim(Boston)[1])
se.hat
```

c. Now estimate the standard error of μ^ using the bootstrap. How does this compare to your answer from (b) ?
```{r}
set.seed(1)
boot.fn <- function(data, index) {
    mu <- mean(data[index])
    return (mu)
}
boot(medv, boot.fn, 1000)
```

c. The bootstrap method standard error of μ^ is similar to the standard error of μ^ in part b when initially estimating the standard error of hat.

d. Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of “medv”. Compare it to the results obtained using t.test(Boston$medv).
```{r}
t.test(medv)
```
```{r}
CI.mu.hat <- c(22.53 - 2 * 0.4119, 22.53 + 2 * 0.4119)
CI.mu.hat
```
d. The confidence interval of the bootstrap method is similar in value to the test function.

e. Based on this data set, provide an estimate, μ^med, for the median value of “medv” in the population.
```{r}
med.hat <- median(medv)
med.hat
```

f. We now would like to estimate the standard error of μ^med. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.
```{r}
boot.fn <- function(data, index) {
    mu <- median(data[index])
    return (mu)
}
boot(medv, boot.fn, 1000)
```

f. The standard error of μ^med is the same median value of 21.2 in both part e and f with an estimated standard error of 0.377.

g. Based on this data set, provide an estimate for the tenth percentile of “medv” in Boston suburbs. Call this quantity μ^0.1.
```{r}
percent.hat <- quantile(medv, c(0.1))
percent.hat
```

h. Use the bootstrap to estimate the standard error of μ^0.1. Comment on your findings.
```{r}
boot.fn <- function(data, index) {
    mu <- quantile(data[index], c(0.1))
    return (mu)
}
boot(medv, boot.fn, 1000)
```

h. The standard error tenth percentile of hat is 12.75 which is also equal to the tenth percentile value of hat in part g. The standard error is 0.492.