---
title: "Data Wrangling"
author: "Ashwin Malshe"
date: "9/3/2019"
output: html_document
---

Install package `pacman` by typing 

install.packages("pacman")

in your console

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)

p_load(dplyr, haven, readxl, nycflights13)

```

`nycflights13` package has flights out of NYC in 2013


Read the data

```{r}
data("flights")
```

Take a peek at the data

```{r}
head(flights)
```
```{r}
tail(flights)
```


## Select variables

Create a new data frame with `month`, `day`, `origin`, `dest`, `air_time`, and `distance`

```{r}
flights2 <- select(flights, month, day, origin, dest, air_time, distance)
```

```{r}
head(flights2)
```



## Diversion: `magrittr` pipe

```{r}
sum(flights2$month)

```


Remove `flights2`
```{r}
rm(flights2)
```


```{r}
flights2 <- flights %>% 
  select(month, day, origin, dest, air_time, distance)
```


## Create new variable

Create a new data set using `flights2` with a new variable called `org_dest` that holds `origin` - `dest` pair. So if origin is `EWR` and destination is `IAH` then `org_dest` will be `EWR_IAH`

```{r}
flights3 <- flights2 %>% 
  mutate(org_dest = paste(origin, dest, sep = "_"),
         distance_km = 1.6 * distance)
```

```{r}
head(flights3)
```


## Filter observations

Create a new data set using `flights3` where you keep observations from January to June.

```{r}
flights4 <- flights3 %>% 
  filter(month <= 6)
```

```{r}
dim(flights3)
dim(flights4)
```

## Sort observations

Sort the dataset `flights4` by increasing `distance` and save it as `flights5`

```{r}
flights5 <- flights4 %>% 
  arrange(-distance, air_time)
```

```{r}
head(flights5)
```


## Use pipe to do all the above steps without creating intermediate data frames

```{r}
flights6 <- flights %>% 
  select(month, day, origin, dest, air_time, distance) %>% 
  mutate(org_dest = paste(origin, dest, sep = "_"),
         distance_km = 1.6 * distance) %>% 
  filter(month <= 6) %>% 
  arrange(distance)
  
```

```{r}
flights7 <- arrange(
  filter(
    mutate(
      select(flights, month, day, origin, dest, air_time, distance),
      org_dest = paste(origin, dest, sep = "_"), 
      distance_km = 1.6 * distance), 
    month <= 6),
  distance)
```

```{r}
identical(flights6, flights7)
```


## Summarize

```{r}
summary(flights6)
```

```{r}
distance_mean <- mean(flights6$distance)
print(distance_mean)

distance_sd <- sd(flights6$distance)
print(distance_sd)

```



Get the mean, median, and standard deviation of `air_time` and `distance` from `flights6`. Print the output rather than saving it.

```{r}

flights6 %>% 
  summarize(air_time_mean = mean(air_time, na.rm = TRUE),
            air_time_sd = sd(air_time, na.rm = TRUE),
            air_time_med = median(air_time, na.rm = TRUE),
            distance_mean = mean(distance),
            distance_sd = sd(distance),
            distance_median = median(distance))

```

## Group and summarize

Get the mean, median, and standard deviation of `air_time` and `distance` from `flights6` for each origin airport. Print the output rather than saving it.

```{r}
flights6 %>% 
  group_by(origin) %>% 
  summarize(air_time_mean = mean(air_time, na.rm = TRUE),
            air_time_sd = sd(air_time, na.rm = TRUE),
            air_time_med = median(air_time, na.rm = TRUE),
            distance_mean = mean(distance),
            distance_sd = sd(distance),
            distance_median = median(distance))

```


## Importing external data files

Import Apple data files

```{r}

```




